Comparing tool variability and numerical variability in fMRI analyses

### 2021-12-01 paper/paper.tex
Edits in methods and results, rewrote discussion.

@@

We replicated the fMRI analysis described as study `ds000001'
in~\cite{schonberg2012decreasing}, relying on the data publicly available
in OpenNeuro at \url{https://openneuro.org/datasets/ds000001} and using the
three main software packages for fMRI data processing, namely FMRIB
Software Library (FSL)~\cite{jenkinson2012fsl}, Analysis of Functional
NeuroImages (AFNI)~\cite{cox1996afni}, and Statistical Parametric
Mapping (SPM)~\cite{penny2011statistical}. We selected this dataset because
comparable analysis pipelines implemented in FSL, AFNI and SPM were already 
publicly available and extensively described in~\cite{bowring2019exploring}.
Furthermore, the work in~\cite{bowring2019exploring} already evaluated the
effect of tool variability for this dataset, which we intended to
complement with the present quantification of numerical variability.

In the selected study, 16 healthy adult subjects participated in the
balloon analog risk task~\cite{lejuez2002evaluation} to measure risk-taking
behavior over three scanning sessions~\cite{schonberg2012decreasing}. We
reused the preprocessing, first-level, and second-level analyses
implemented by~\cite{bowring2019exploring} consistently across all three
software packages. Table~\ref{table:pipeline-steps} summarizes the analytical steps in each
pipeline. More details are available in~\cite{bowring2019exploring}.

@@

To introduce controlled amounts of numerical noise in the analyses, we used
the Fuzzy Libmath library~\cite{salari2021accurate}, a version of the GNU
mathematical library (libmath) instrumented with Monte-Carlo arithmetic.
Monte-Carlo arithmetic simulates roundoff and catastrophic cancellation
errors by introducing a controlled amount of noise in the floating-point
operations through the following perturbation model~\cite{Parker1997-qq}:

\begin{equation} \label{eq:mca_inexact}
  inexact(x) = x + 2^{e_x-t}\xi
\end{equation}

where $e_x$ is the exponent in the floating-point representation of $x$,
$t$ is the virtual precision --- the number of unperturbed bits in the
mantissa of $x$ ---, and $\xi$ is a random uniform variable of
$(-\frac{1}{2}, \frac{1}{2})$. The perturbation is introduced using 
Verificarlo~\cite{denis2015verificarlo}, an LLVM compiler supporting Monte-Carlo 
arithmetic and other types of numerical analyses.

The instrumented libmath functions are loaded in the pipeline using
LD\_PRELOAD, a Linux mechanism to force load a shared library into an
executable. This allows functions defined in Fuzzy Libmath to transparently
overload the original ones without the need to modify or recompile the
analysis pipeline.

Fuzzy Libmath only introduces perturbations in the output values of
mathematical functions but not in their input values or within their
implementation. This is done by wrapping the original functions and
applying function \texttt{inexact} to their returned values.
Listing~\ref{algo:wrapper} shows an example of this wrapping for the
\texttt{log} function in single and double precision. In this wrapper, the
original function is called through \texttt{dlsym}, a function that returns
the memory address of a symbol --- in our case \texttt{RTLD\_NEXT}, the
address of the next occurrence of the function in memory. Function wrappers
are compiled with Verificarlo. Verificarlo instruments the result of the
addition between the original function output and a floating-point zero.

@@

To validate instrumentations, we verified that two consecutive IEEE executions are identical;
there are no randomness effects in all pipelines.
We also listed all the pipeline library dependencies using the ldd Linux utility.
We ensured that (1) all pipelines are dynamically linked to the libmath library
(2) there is no alternative implementation for the basic mathematical functions in these dependencies.
Finally, we verified that Fuzzy libmath affects computational results.

@@

We measured between-tool (BT) variability by running the pipelines
described in~\cite{bowring2019exploring} with FSL version 5.0.10, AFNI
version 18.1.09, and SPM12 version r7771 executed with GNU/Octave version
5.2. All these software versions were identical to the study
in~\cite{bowring2019exploring} except for SPM for which we used Octave
instead of MATLAB to enable mathematical function instrumentation using
Fuzzy Libmath. Indeed, MATLAB uses its own built-in mathematical functions,
which prevents the use of Fuzzy Libmath. In AFNI, we set the number of
threads to 7 even though AFNI executions
in~\cite{bowring2019exploring} were single-threaded. This was meant to
reduce the time overhead resulting from Fuzzy Libmath instrumentation. All
the analyses were conducted on the CentOS 7.3 operating system. The
computations were performed on \href{https://www.computecanada.ca}{Compute
Canada's} Béluga cluster nodes, each with 2$\times$ Intel Gold 6148 Skylake
@ 2.4 GHz (40 cores/node) CPU and 8~GB of RAM per core. We encapsulated the
above-mentioned software packages in Docker container images available at
\url{https://github.com/big-data-lab-team/fuzzy-neurotools}.

We measured within-tool (WT) variability by running the same analyses three
times using Fuzzy Libmath with a virtual precision of $t=53$~bits for
double-precision values and $t=24$~bits for single-precision values. These
values were chosen such that the numerical perturbation simulates machine
error, the relative error between a real number and its floating-point
representation. The resulting samples are equally plausible estimates of
the true numerical result. Moreover, we repeated the FSL analyses for virtual
precisions ranging from $t=1$~bit to $t=24$~bits for single-precision
and double-precision values. We also repeated the FSL analyses for virtual 
precisions ranging from $t=24$~bits to $t=53$~bits for double-precision values, having 
set the virtual precision to $t=24$~bits for single-precision values.

We evaluated WT and BT variability for thresholded as well as unthresholded
group-level and subject-level t-statistics maps, by computing the
standard deviation of t-statistic maps. For BT, we computed the standard deviation 
across a given pair of tools.
For WT, we computed the standard deviation across 
Fuzzy Libmath samples of a given tool. Moreover, we computed WT for a pair of tools (A, B) as follows:
\begin{equation}
  \sigma_{WT(A,B)}^2 = \sigma_{WT(A)}^2 + \sigma_{WT(B)}^2,
  \label{eq:wt-pair}
\end{equation}
where $WT(.)$ is the within-tool variability for a given tool.

Further, from the thresholded maps, we determined regional instability 
between activation clusters in the 360 regions in the Human
Connectome Project Multi-Modal Parcellation atlas version 1.0
(HCP-MMP1.0)~\cite{glasser2016multi}. For BT, a region was unstable if it had different activation behavior between pair of tools.
For WT, we considered a region unstable if it had a blinking activation between Fuzzy samples of a given tool.

@@

We verified the correctness of our analyses by comparing our unperturbed
t-statistic group maps with the ones obtained
in~\cite{bowring2019exploring}. For FSL, we found
identical checksums. For SPM, the checksums were different but differences
were visually imperceptible. For AFNI, the activation maps were similar
overall, however, differences were more noticeable visually. 
Overall, our results correctly reproduced
the ones presented in~\cite{bowring2019exploring}. The observed
differences were small and might be due to the use of Octave vs MATLAB in
SPM or different hardware. We performed visual quality control of the AFNI and SPM results
for each individual subject.
\tristan{add a figure to show the differences for SPM (Octave) and AFNI}

@@

Summary statistics for the group-level unthresholded and thresholded
t-statistics are reported in Table~\ref{table:pipeline-stats}.
Overall, BT variability was larger than WT variability in both thresholded and unthresholded maps.
This was confirmed by computing the T-test and Wilcoxon signed-rank statistics between BT and WT as:

\begin{equation*}
  \left\{
    \begin{aligned}
      & BT(a, b), WT(a),\\
      & BT(a, b), WT(b)
    \end{aligned}  
    \quad  a,b \in [FSL, SPM, AFNI] \;
  \right\}
  \label{eq:t-tests}
\end{equation*}

We obtained statistically significant values between BT and WT
variability with t-value $> 70$ and p-value $< 0.05$.

%%%%%%%%%% Summary of statstics %%%%%%%%
\setlength{\tabcolsep}{5pt}
\begin{table}[h]
    \centering
    \begin{tabular}{cccc|cc}
        \toprule
        \multirow{2}{*}{}& {} & \multicolumn{2}{c}{Thresholded} & \multicolumn{2}{c}{Unthresholded} \\
        \cmidrule{3-4} \cmidrule{5-6} \\
        {} & {} & Mean & Std. dev. & Mean & Std. dev. \\
        \midrule
        \rowcolor{lightgray}
        {Between Tools} & FSL vs. SPM        &  1.282       & 0.525      & 0.443     & 0.344  \\
        \rowcolor{lightgray}
        {(BT)} & FSL vs. AFNI                &  1.548       & 0.616      & 0.547     & 0.441  \\
        \rowcolor{lightgray}
        {} & AFNI vs. SPM                    &  1.475       & 0.672      & 0.608     & 0.477  \\
        {Within Tool} & FSL                  &  0.354       & 0.491      & 0.082     & 0.065  \\
        {(WT)}   & SPM                       &  0.252       & 0.448      & 0.054     & 0.045  \\
        {}   & AFNI                          &  0.434       & 0.524      & 0.128     & 0.135  \\
        \bottomrule
    \end{tabular}
    \caption{Summary statistics of voxel-wise standard deviations
    in group-level t-statistics in BT and WT.}
    \label{table:pipeline-stats}
\end{table}

@@

Figure~\ref{fig:unthresh-maps} compares t-statistics standard deviation maps
across different pipelines (BT variability) and across Fuzzy Libmath
samples (WT variability). While WT variability remains lower than BT
variability overall, it approaches it in some regions, shown in white in
Figure~\ref{fig:unthresh-maps}-\textbf{C}.

Interestingly, a group of voxels \tristan{...}
BT and WT variability appear correlated in some regions
(Figure~\ref{fig:unthresh-correlation}).

%%%%%%%%%% Var. of Unthresh %%%%%%%%
\begin{figure*}[ht]
    \fbox{\begin{minipage}{\dimexpr \textwidth-2\fboxsep-2\fboxrule}
      \centering
      \includegraphics[width=\textwidth]{figures/std/gl-unthresh.png}
      %\caption{Standard deviation of thresholded t-statistics map on template surface}
    \caption{Unthresholded group t-statistics standard deviations computed between tools (\textbf{A}) and within tools (\textbf{B}), and difference between them (\textbf{C}).}
    \label{fig:unthresh-maps}
    \end{minipage}}
  \end{figure*}

  %%%%%%%%%% Corr. plot of tstats%%%%%%%%
  \begin{figure*}[ht]
    \fbox{\begin{minipage}{\dimexpr \textwidth-2\fboxsep-2\fboxrule}
      \begin{subfigure}[ht]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/std-corr-unthresh-plot.png}
        %\caption{Standard deviation of thresholded t-statistics map on template surface}
      \end{subfigure}
      \hfill
      \begin{subfigure}[ht]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/std/correlated-unthresh.png}
        %\caption{Standard deviation of thresholded t-statistics map on template surface}
      \end{subfigure}
      \caption{
        Voxel-wise comparison of BT and WT variabilty. The bottom row represents 
        voxels where the BT/WT ratio is in (0.5, 2).}
    \label{fig:unthresh-correlation}
    \end{minipage}}
  \end{figure*}

@@

  Summary statistics for subject-level unthresholded t-statistics maps are
  reported in Table~\ref{table:unthresh-maps-subjects}. Overall, results
  represent similar findings with group-level t-statistics between WT and
  BT variations.

  Figure~\ref{fig:unthresh-maps-sbj} shows BT and WT variability of the
  unthresholded t-statistics for the subject with the highest WT
  variability among the 16 subjects.
  For this particular subject, we observed substantial WT variations with the average standard deviation 0.099 compared to 
  the average standard deviation 0.466 for BT variations.
  As can be seen in the Figure, WT variability approaches and even surpasses BT variability in some regions.
  This subject also revealed that WT variability mainly concentrated in some regions  
  contrary to the corresponding variability from unthresholded group-level t-statistics which was uniformly distributed across the brain.

  %%%%%%%%%% Summary of statstics %%%%%%%%
\setlength{\tabcolsep}{5pt}
\begin{table}[h]
    \centering
    \begin{tabular}{cccc}
        \toprule
        \multirow{2}{*}{}& {} & \multicolumn{2}{c}{Unthresholded} \\
        \cmidrule{3-4} \\
        {} & {} & Mean & Std. dev. \\
        \midrule
        \rowcolor{lightgray}
        {Between Tools} & FSL vs. SPM        &  0.366       & 0.293     \\
        \rowcolor{lightgray}
        {(BT)} & FSL vs. AFNI                &  0.439       & 0.352     \\
        \rowcolor{lightgray}
        {} & AFNI vs. SPM                    &  0.491       & 0.381     \\
        {Within Tool} & FSL                  &  0.077       & 0.054     \\
        {(WT)}   & SPM                       &  0.048       & 0.037     \\
        {}   & AFNI                          &  0.108       & 0.131     \\
        \bottomrule
    \end{tabular} 
    \caption{Summary statistics of voxel-wise standard deviations in subject-level t-statistics in BT and WT.
    The voxel-wise standard deviations were averaged across 16 subjects.}
    \label{table:unthresh-maps-subjects}
\end{table}

  %%%%%%%%%% Var. of Unthresh sbj05%%%%%%%%
\begin{figure*}[ht]
  \fbox{\begin{minipage}{\dimexpr \textwidth-2\fboxsep-2\fboxrule}
    \centering
    \includegraphics[width=\textwidth]{figures/std/sbj05-std.png}
    %\caption{Standard deviation of thresholded t-statistics map on template surface}
    \caption{Unthresholded subject t-statistics standard deviations computed between tools (\textbf{A}) and within tools (\textbf{B}), and difference between them (\textbf{C}).
    This is for the subject with the highest WT variability.}
  \label{fig:unthresh-maps-sbj}
  \end{minipage}}
\end{figure*}

\subsection{WT variability across virtual precisions}

While the previous results were obtained at the virtual precision of 53~bits for double-precision floating-point values
and 24~bits for single-precision values, we evaluated WT variability across different virtual precisions for FSL.
Figure~\ref{fig:across-precisions} represents the root-mean-square error (RMSE) between the standard deviations per voxel
in BT and WT resulted from the Fuzzy FSL,
substantiating that how variabilities in both conditions are close to each other across different virtual precisions.

We determined the virtual precision of t=17 bits as the precision that minimized the RMSE between BT and WT on average.
This is the precision that numerical perturbations in FSL more closely simulated BT variability.
The brain maps of standard deviations between BT and WT obtained from the Fuzzy FSL at t=17 bits (Figure~\ref{fig:gnp-mni})
show significant WT variability, particularly for some regions in the frontal, parietal, and limbic lobes, close to BT variability.

  %%%%%%%%%% plot different precisions%%%%%%%%
  \begin{figure}[ht]
    \fbox{\begin{minipage}{\dimexpr \columnwidth-2\fboxsep-2\fboxrule}
        \centering
        \includegraphics[width=\columnwidth]{figures/rmse-precisions.png}
        %\caption{Standard deviation of thresholded t-statistics map on template surface}
      \caption{Voxel-wise RMSE between BT and WT variability maps for different virtual precisions. WT is computed from three samples of FSL tool.}
    \label{fig:across-precisions}
    \end{minipage}}
  \end{figure}

  %%%%%%%%%% plot different precisions%%%%%%%%
  \begin{figure}[ht]
    \fbox{\begin{minipage}{\dimexpr \textwidth-2\fboxsep-2\fboxrule}
        \centering
        \includegraphics[width=\textwidth]{figures/bg_global_precision.pdf}
        %\caption{Standard deviation of thresholded t-statistics map on template surface}
        \caption{Unthresholded group t-statistics standard deviations computed between tools (\textbf{A}) and within tools (\textbf{B}), and difference between them (\textbf{C}).
        WT is computed from three samples of FSL tool at the virtual precision of t=17 bits.}
    \label{fig:gnp-mni}
    \end{minipage}}
  \end{figure}

@@

  Similar conclusions are drawn from the unthresholded t-statistics maps
  (Figure~\ref{fig:thresh-maps}).
  Figure~\ref{fig:cm-thresh} compares instability of regional activations,
  showing that most of the regions have a similar activation pattern in both BT and WT, mostly are stable.
  However, some regions have a different activation behavior, mainly unstable in BT and stable in WT.
  % showing more severity of BT instability than WT.

  %%%%%%%%%% Var. of Thresh %%%%%%%%
  \begin{figure*}[ht]
      \fbox{\begin{minipage}{\dimexpr \textwidth-2\fboxsep-2\fboxrule}
        \centering
        \includegraphics[width=\textwidth]{figures/std/gl-thresh.png}
      \caption{Thresholded group t-statistics standard deviation maps
      computed across tools (\textbf{A}) and Fuzzy Libmath samples
      (\textbf{B}) as defined in Eq~\ref{eq:wt-pair}.}
      \label{fig:thresh-maps}
      \end{minipage}}
    \end{figure*}

    % %%%%%%%%%% Dice plot of thresholded tstats%%%%%%%%
    \begin{figure*}[ht]
      \centering
      \includegraphics[width=\textwidth]{figures/dice_analysis-plot.png}
      \caption{Heatmap of confusion matrices for instability of regional activations
      in BT and WT using the HCP-MMP1.0 parcellation of ~\cite{glasser2016multi}. 
      A region is unstable (=1) if it has different activation behavior in both tools
      for BT or different fuzzy samples in each tool for WT; otherwise, it is stable (=0).}
      \label{fig:cm-thresh}
    \end{figure*}

@@

We represented differences in the results of an fMRI analysis between numerical variability
and the variability arisen from software package variations.
Although BT variability was larger than WT variability on average,
we demonstrated regions in the brain maps that numerical perturbations could simulate BT variability.
The correlation of these variabilities suggests that numerical perturbations can be used to evaluate the robustness of the analyses
to inter-software disparities.

While switching between \fslspm was generated the least BT and WT variabilities consistently, the most level of these variabilities was identified
between \fslafni and \afnispm. %We also found AFNI as the most sensitive tool to numerical perturbations.
These results do not indicate which software package is better or worse since there is no ground-truth model for the analyses.
However, this study revealed the importance of numerical instability in the selected tools for doing fMRI analyses.
It provided a view of related numerical variability of which should be taken into consideration by the software developers.

>>>>

\tristan{We investigate two questions: (1) how do numerical and between-tool variability compare? (2) are they correlated?}

@@

We replicated the fMRI analysis described as study `ds000001'
in~\cite{schonberg2012decreasing}, relying on the data publicly available
in OpenNeuro at \url{https://openneuro.org/datasets/ds000001} and using
three widely-used software packages for fMRI data processing, namely FMRIB
Software Library (FSL)~\cite{jenkinson2012fsl}, Analysis of Functional
NeuroImages (AFNI)~\cite{cox1996afni}, and Statistical Parametric
Mapping (SPM)~\cite{penny2011statistical}. We selected this dataset because
comparable analysis pipelines implemented in FSL, AFNI and SPM were already 
publicly available and extensively described in~\cite{bowring2019exploring}.
Furthermore, the work in~\cite{bowring2019exploring} already evaluated the
effect of tool variability for this dataset, which we intended to
complement with the present quantification of numerical variability.

In the selected study, 16 healthy adult subjects participated in the
balloon analog risk task~\cite{lejuez2002evaluation} to measure risk-taking
behavior over three scanning sessions~\cite{schonberg2012decreasing}. We
reused the preprocessing, first-level, and second-level analyses
implemented by~\cite{bowring2019exploring} consistently across all three
software packages. Table~\ref{table:pipeline-steps} summarizes the analytical steps in each
pipeline.

@@

To introduce numerical noise in the analyses, we used
the Fuzzy Libmath library~\cite{salari2021accurate}, a version of the GNU
mathematical library (libmath) instrumented with Monte-Carlo arithmetic.
Monte-Carlo arithmetic simulates roundoff and catastrophic cancellation
errors by introducing a controlled amount of noise in the floating-point
operations through the following perturbation~\cite{Parker1997-qq}:

\begin{equation} \label{eq:mca_inexact}
  inexact(x) = x + 2^{e_x-t}\xi,
\end{equation}
where $e_x$ is the exponent in the floating-point representation of $x$,
$t$ is the virtual precision (the number of unperturbed bits in the
mantissa of $x$), and $\xi$ is a random uniform variable of
$(-\frac{1}{2}, \frac{1}{2})$. We introduced the perturbation using 
Verificarlo~\cite{denis2015verificarlo}, an LLVM compiler supporting Monte-Carlo 
arithmetic and other types of numerical instrumentations.

We loaded the instrumented libmath functions in the pipeline using
LD\_PRELOAD, a Linux mechanism to force-load a shared library into an
executable. This mechanism allows functions defined in Fuzzy Libmath to transparently
overload the original ones without the need to modify or recompile the
analysis pipeline.

Fuzzy Libmath introduces numerical perturbations in the values returned by
mathematical functions but not in their input values or within their
implementation. This is done by wrapping the original functions and
applying function \texttt{inexact} to their returned values.
Listing~\ref{algo:wrapper} shows an example of this wrapping for the
\texttt{log} function in single and double precision. In this wrapper, the
original function is called through \texttt{dlsym}, a function that returns
the memory address of a symbol --- in our case \texttt{RTLD\_NEXT}, the
address of the next occurrence of the function in memory. Function wrappers
are compiled with Verificarlo. Verificarlo instruments the result of the
addition between the original function output and a floating-point zero.

@@

To validate our instrumentations, we first verified that non-instrumented
executions of the same pipeline on the same dataset led to identical
results \tristan{did you fix random seeds and if so in which pipelines?}. We also listed the pipeline
library dependencies using the \texttt{ldd} Linux utility and verified that
(1) the tested pipelines were dynamically linked to the GNU libmath library, and 
(2) there was no alternative implementation of elementary mathematical functions in the pipeline dependencies.
Finally, we verified that the use of Fuzzy libmath affected computational results.

@@

We measured between-tool (BT) variability by running the pipelines
described in~\cite{bowring2019exploring} with FSL version 5.0.10, AFNI
version 18.1.09, and SPM12 version r7771 executed with GNU/Octave version
5.2. These software versions were identical to the study
in~\cite{bowring2019exploring} except for SPM for which we used GNU/Octave
instead of MATLAB to enable mathematical function instrumentation using
Fuzzy Libmath. Indeed, MATLAB uses its own built-in mathematical functions,
which prevents the use of Fuzzy Libmath. In AFNI, we set the number of
threads to 7 even though AFNI executions
in~\cite{bowring2019exploring} were single-threaded. This was meant to
reduce the time overhead resulting from Fuzzy Libmath instrumentation. All
the analyses were conducted on the CentOS 7.3 operating system. The
computations were performed on \href{https://www.computecanada.ca}{Compute
Canada's} Béluga cluster nodes, each with 2$\times$ Intel Gold 6148 Skylake
@ 2.4 GHz (40 cores/node) CPU and 8~GB of RAM per core. We encapsulated the
above-mentioned software packages in Docker container images available at
\url{https://github.com/big-data-lab-team/fuzzy-neurotools}.

We measured within-tool (WT) variability by running the same analyses three
times using Fuzzy Libmath with a virtual precision of $t=53$~bits for
double-precision values and $t=24$~bits for single-precision values. These
values were chosen such that the numerical perturbation simulates machine
error, the minimal distance between two consecutive floating-point numbers. The resulting samples are equally plausible estimates of
the true numerical result. Moreover, to evaluate numerical perturbations at different magnitudes, 
we repeated the FSL analyses for virtual
precisions ranging from $t=1$~bit to $t=24$~bits for single-precision
and double-precision values. We also repeated the FSL analyses for virtual 
precisions ranging from $t=24$~bits to $t=53$~bits for double-precision values, having 
set the virtual precision to $t=24$~bits for single-precision values.

We evaluated WT and BT variability for thresholded as well as unthresholded
group-level and subject-level t-statistics maps, by computing the
standard deviation of t-statistic maps across pipelines (BT) or Fuzzy Libmath samples (WT). For BT, we computed the standard deviation 
across a given pair of tools.
For WT, we computed the standard deviation across the 3 
Fuzzy Libmath samples of a given tool. Moreover, we computed WT variability for a pair of tools (A, B) as follows:
\begin{equation}
  \sigma_{WT(A,B)}^2 = \sigma_{WT(A)}^2 + \sigma_{WT(B)}^2,
  \label{eq:wt-pair}
\end{equation}
where $\sigma_{WT(.)}$ is the within-tool variability for a given tool.

Further, from the thresholded maps, we determined regional instability
between activation clusters in the 360 regions in the Human Connectome
Project Multi-Modal Parcellation atlas version 1.0
(HCP-MMP1.0)~\cite{glasser2016multi}. For BT, we considered a region
unstable for a pair of tools if it contained activated voxels for a tool
but not for the other one. For WT, we considered a region unstable for a
pair of tools (A, B) if for tool A or tool B it contained activated voxels only for some Fuzzy Libmath
samples.

@@

We verified the correctness of our analyses by comparing our unperturbed
t-statistic group maps with the ones obtained
in~\cite{bowring2019exploring}. For FSL, we found
identical checksums. For SPM, the checksums were different but differences
were visually unnoticeable. For AFNI, the activation maps were similar
overall, however, differences were more noticeable visually. 
The observed
differences remained small and might be due to the use of Octave vs MATLAB in
SPM. We performed visual quality control of the AFNI and SPM results
for each individual subject \tristan{and confirmed that ...}.
\tristan{add a figure to show the differences for SPM (Octave) and AFNI}

@@

Summary statistics for the group-level unthresholded and thresholded
t-statistics are reported in Table~\ref{table:pipeline-stats}. BT
variability for tool pair (A, B) was larger than WT variability for tool A and tool B for any tool pair, in both thresholded and
unthresholded maps (Wilcoxon signed-rank test, t-test
p~\textless~$10^{-163}$ for all tests).
% This was confirmed by computing the T-test and Wilcoxon signed-rank statistics between BT and WT as:

% \begin{equation*}
%   \left\{
%     \begin{aligned}
%       & BT(a, b), WT(a),\\
%       & BT(a, b), WT(b)
%     \end{aligned}  
%     \quad  a,b \in [FSL, SPM, AFNI] \;
%   \right\}
%   \label{eq:t-tests}
% \end{equation*}

% We obtained statistically significant values between BT and WT
% variability with t-value $> 70$ and p-value $< 0.05$.

%%%%%%%%%% Summary of statstics %%%%%%%%
\setlength{\tabcolsep}{5pt}
\begin{table}[h]
    \centering
    \begin{tabular}{cccc|cc}
        \toprule
        \multirow{2}{*}{}& {} & \multicolumn{2}{c}{Thresholded} & \multicolumn{2}{c}{Unthresholded} \\
        \cmidrule{3-4} \cmidrule{5-6} \\
        {} & {} & Mean & Std. dev. & Mean & Std. dev. \\
        \midrule
        \rowcolor{lightgray}
        {Between Tools} & FSL vs. SPM        &  1.282       & 0.525      & 0.443     & 0.344  \\
        \rowcolor{lightgray}
        {(BT)} & FSL vs. AFNI                &  1.548       & 0.616      & 0.547     & 0.441  \\
        \rowcolor{lightgray}
        {} & AFNI vs. SPM                    &  1.475       & 0.672      & 0.608     & 0.477  \\
        {Within Tool} & FSL                  &  0.354       & 0.491      & 0.082     & 0.065  \\
        {(WT)}   & SPM                       &  0.252       & 0.448      & 0.054     & 0.045  \\
        {}   & AFNI                          &  0.434       & 0.524      & 0.128     & 0.135  \\
        \bottomrule
    \end{tabular}
    \caption{Voxel-wise mean and standard deviation of BT and WT variability
    in group-level t-statistics maps.}
    \label{table:pipeline-stats}
\end{table}

@@

Figure~\ref{fig:unthresh-maps} compares BT and WT variability in
unthresholded maps. WT variability remains lower than BT variability
overall. Interestingly, BT and WT variability appear correlated in a subset
of voxels scattered across the brain
(Figure~\ref{fig:unthresh-correlation}), but the reason for this apparent
correlation remains unknown.

%%%%%%%%%% Var. of Unthresh %%%%%%%%
\begin{figure*}[ht]
    \fbox{\begin{minipage}{\dimexpr \textwidth-2\fboxsep-2\fboxrule}
      \centering
      \includegraphics[width=\textwidth]{figures/std/gl-unthresh.png}
      %\caption{Standard deviation of thresholded t-statistics map on template surface}
    \caption{Unthresholded group-level variability computed between tools (\textbf{A}) and within tools (\textbf{B}), and difference between them (\textbf{C}).}
    \label{fig:unthresh-maps}
    \end{minipage}}
  \end{figure*}

  %%%%%%%%%% Corr. plot of tstats%%%%%%%%
  \begin{figure*}[ht]
    \fbox{\begin{minipage}{\dimexpr \textwidth-2\fboxsep-2\fboxrule}
      \begin{subfigure}[ht]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/std-corr-unthresh-plot.png}
        %\caption{Standard deviation of thresholded t-statistics map on template surface}
      \end{subfigure}
      \hfill
      \begin{subfigure}[ht]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/std/correlated-unthresh.png}
        %\caption{Standard deviation of thresholded t-statistics map on template surface}
      \end{subfigure}
      \caption{
        Voxel-wise comparison of BT and WT variabilty. The bottom row shows 
        voxels where the BT/WT ratio is in (0.5, 2).}
    \label{fig:unthresh-correlation}
    \end{minipage}}
  \end{figure*}

@@

  Summary statistics for subject-level unthresholded t-statistics maps are
  reported in Table~\ref{table:unthresh-maps-subjects}. As for group-level
  maps, BT variability was consistently larger than WT variability
  \tristan{report statistical tests}.

  Figure~\ref{fig:unthresh-maps-sbj} shows BT and WT variability for
  the subject with the highest WT variability among the 16 subjects. For
  this particular subject, WT variability approaches and even surpasses BT
  variability in some regions. This subject also revealed that WT
  variability mainly concentrated in some regions.

  %%%%%%%%%% Summary of statstics %%%%%%%%
\setlength{\tabcolsep}{5pt}
\begin{table}[h]
    \centering
    \begin{tabular}{cccc}
        \toprule
        \multirow{2}{*}{}& {} & \multicolumn{2}{c}{Unthresholded} \\
        \cmidrule{3-4} \\
        {} & {} & Mean & Std. dev. \\
        \midrule
        \rowcolor{lightgray}
        {Between Tools} & FSL vs. SPM        &  0.366       & 0.293     \\
        \rowcolor{lightgray}
        {(BT)} & FSL vs. AFNI                &  0.439       & 0.352     \\
        \rowcolor{lightgray}
        {} & AFNI vs. SPM                    &  0.491       & 0.381     \\
        {Within Tool} & FSL                  &  0.077       & 0.054     \\
        {(WT)}   & SPM                       &  0.048       & 0.037     \\
        {}   & AFNI                          &  0.108       & 0.131     \\
        \bottomrule
    \end{tabular} 
    \caption{Voxel-wise mean and standard deviation of BT and WT variability
    in subject-level t-statistics maps (averages on 16 subjects).}
    \label{table:unthresh-maps-subjects}
\end{table}

  %%%%%%%%%% Var. of Unthresh sbj05%%%%%%%%
\begin{figure*}[ht]
  \fbox{\begin{minipage}{\dimexpr \textwidth-2\fboxsep-2\fboxrule}
    \centering
    \includegraphics[width=\textwidth]{figures/std/sbj05-std.png}
    %\caption{Standard deviation of thresholded t-statistics map on template surface}
    \caption{For subject with highest WT variability, unthresholded subject-level variability computed between tools (\textbf{A}) and within tools (\textbf{B}), and difference between them (\textbf{C}).}
  \label{fig:unthresh-maps-sbj}
  \end{minipage}}
\end{figure*}

\subsection{Group-level untresholded maps at different virtual precisions}

While the previous results were obtained at the virtual precision of 53~bits for double-precision values
and 24~bits for single-precision values, we also evaluated WT variability across different virtual precisions for FSL.
Figure~\ref{fig:across-precisions} shows the root-mean-square error (RMSE) between BT and WT variability maps.
The virtual precision of t=17 bits minimized the RMSE between BT and WT.
Figure~\ref{fig:gnp-mni} shows the corresponding BT and WT maps obtained at t=17 bits, showing 
that BT and WT reach comparable magnitudes in some regions.

  %%%%%%%%%% plot different precisions%%%%%%%%
  \begin{figure}[ht]
    \fbox{\begin{minipage}{\dimexpr \columnwidth-2\fboxsep-2\fboxrule}
        \centering
        \includegraphics[width=\columnwidth]{figures/rmse-precisions.png}
        %\caption{Standard deviation of thresholded t-statistics map on template surface}
      \caption{Voxel-wise RMSE between BT and WT variability maps for different virtual precisions. \tristan{Remove (std) from y label}}
    \label{fig:across-precisions}
    \end{minipage}}
  \end{figure}

  %%%%%%%%%% plot different precisions%%%%%%%%
  \begin{figure}[ht]
    \fbox{\begin{minipage}{\dimexpr \textwidth-2\fboxsep-2\fboxrule}
        \centering
        \includegraphics[width=\textwidth]{figures/bg_global_precision.pdf}
        %\caption{Standard deviation of thresholded t-statistics map on template surface}
        \caption{Unthresholded group t-statistics standard deviations computed between tools (\textbf{A}) and within tools (\textbf{B}), and difference between them (\textbf{C}).
        WT is computed at the virtual precision of t=17 bits. \tristan{remove (AFNI,SPM)}}
    \label{fig:gnp-mni}
    \end{minipage}}
  \end{figure}

@@

  Figure~\ref{fig:thresh-maps} compares BT and WT for thresholded group-level maps.
  While BT is consistently larger than WT, thresholding is an unstable operation 
  that results in comparable magnitudes for BT and WT within activation clusters where BT variability is low.

  Figure~\ref{fig:cm-thresh} compares regional activation instability for BT and WT. The average ratio
  of unstable regions was 26\% for BT and 9.2\% for WT. Moreover, the average F1 score was 0.28. Therefore, regional 
  WT instability does not appear to be a 
  good predictor of BT instability.

  %%%%%%%%%% Var. of Thresh %%%%%%%%
  \begin{figure*}[ht]
      \fbox{\begin{minipage}{\dimexpr \textwidth-2\fboxsep-2\fboxrule}
        \centering
        \includegraphics[width=\textwidth]{figures/std/gl-thresh.png}
      \caption{Thresholded group t-statistics standard deviation maps
      computed across tools (\textbf{A}) and Fuzzy Libmath samples
      (\textbf{B}).}
      \label{fig:thresh-maps}
      \end{minipage}}
    \end{figure*}

    % %%%%%%%%%% Dice plot of thresholded tstats%%%%%%%%
    \begin{figure*}[ht]
      \fbox{\begin{minipage}{\dimexpr \textwidth-2\fboxsep-2\fboxrule}
      \centering
      \includegraphics[width=\textwidth]{figures/dice_analysis-plot.png}
      \caption{Confusion matrices of activation instability
      in BT and WT among the 360 regions from the HCP-MMP1.0 parcellation. \tristan{plot a single color map for 3 heatmaps}}
      \label{fig:cm-thresh}
    \end{minipage}}
    \end{figure*}

@@

In fMRI \emph{group analyses}, machine error remains an order of magnitude
smaller than between-tool variability. Group analyses have a regularization
effect toward numerical noise, which is expected to amplify as sample size
increases. Therefore, for fMRI studies with large sample sizes, machine
error can safely be neglected with respect to between-tool variability. The
recommendation made in~\cite{botvinik2020variability} to rely on ``multiverse" analyses where
multiple analysis tools are compared is therefore likely to successfully correct
for machine error in group studies.

However, machine error remains important in group analyses that are based
on a single tool, as is commonly the case in current fMRI studies. In
particular, in our study, the inherent instability of thresholding was
triggered by machine error in 9\% of 360 brain regions, which indicates
that machine error might have impacted neuroscientific conclusions related
to these regions.

In \emph{subject analyses}, machine error and between-tool variabilty can
become comparable for some subjects in some regions. This observation is particularly
relevant to the development of fMRI-based biomarkers aiming at
individualized phenotype predictions. Machine error is expected to play
a non-negligible role in such analyses, even when predictions combine
results produced by multiple tools.
